Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-01-27 paper/ijcai16.tex
paper: modificato abstract e esperimenti

\begin{abstract}
In this paper we propose an approach to preference elicitation that is suitable to large
configuration problem domains where stated preferences can be noisy.
Our setwise max-margin method can be viewed as a generalization of max-margin learning to sets,
and can produce a set of ``diverse'' items that can be used to ask informative queries to the user.
Moreover, the approach can encourage sparsity in the parameter space, in order to favor the assessment of utility towards combinations of weights that concentrate on just few features.
We present an optimization formulation and show, with simulations, how our approach is able to efficiently elicit the user's utility function even in very large datasets.

\end{abstract}

@@

In Figure~\ref{fig:} we plot, for each setting, the behavior of the utility
loss:
%
$$ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right) $$
%
varying the number of collected user responses. Here $u(\cdot)$ is the
true unknown user utility, and $\vx^*$ is the solution recommended to
the user after the elicitation phase (see Algorithm~\ref{alg:setmargin}).

\begin{itemize}

    \item \ref{fig:comparison} setmargin vs Bayes. similar loss wrt Bayes.
    variance slighly higher on synthetic 4. funzioniamo meglio su sparso al
    contrario degli altri. Guo e' piantato al primo utente da tre giorni
    su synthetic 5.  Paolo procede su S-5 da qualche ora (~8) (1290s su
    synthetic-5 uniforme per completamento di un utente).

    \stefano{@Paolo: varianza tempi?}

    \item \ref{fig:selfcomparison} crossval ogni 5 iterazioni, il che vuol
    dire diverso numero di queri per $k$ diversi.

\end{itemize}

\paragraph{Cognitive effort.} Contrarily to most other preference elicitation
methods, which generate a single candidate option per iteration, {\sc
SetMargin} constructs a whole {\em set} of options. We thus can not leverage
standard query selection strategies \stefano{which ones?}.

The simplest solution is to query the user about all $k \choose 2$ pairs of
configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
number of queries does not grow large in practice.  An appealing alternative
involves using a modified sorting algorithm, such as merge-sort, to order the
set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
approach is very susceptible to noise: a single inconsistent user response may
affect all successive comparisons, leading to a cascading effect. While more
robust strategies may be conceived, empirically we found that the simpler
quadratic scheme tends to work reasonably well.

\andrea{rewrite as footnote}

>>>>

\begin{abstract}
  In this paper we propose an approach to preference elicitation that
  is suitable to large configuration spaces beyond the reach of
  existing state-of-the-art approaches. Our setwise max-margin method
  can be viewed as a generalization of max-margin learning to sets,
  and can produce a set of ``diverse'' items that can be used to ask
  informative queries to the user.  Moreover, the approach can
  encourage sparsity in the parameter space, in order to favor the
  assessment of utility towards combinations of weights that
  concentrate on just few features.  We present a mixed integer linear
  programming formulation and show how our approach compares
  favourably with Bayesian preference elicitation alternatives and
  easily scales to realistic datasets.
\end{abstract}

@@

\andrea{give names to competitors}

In Figure~\ref{fig:comparison} we report solution quality and timing
values for increasing number of collected user responses, for the
different competitors on each of the four different utility vector
distributions. Solution quality is measured in terms of utility loss:
%
$$ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right) $$
%
where $u(\cdot)$ is the true unknown user utility, and $\vx^*$ is the
solution recommended to the user after the elicitation phase (see
Algorithm~\ref{alg:setmargin}). Computational cost is measured in
terms of cumulative time. Given that Guo and Viappiani are
single-threaded, we disabled multi-threading when running our
algorithm in these comparisons. All experiments were run on a XXXXXX.
In all figures, thick lines indicate median values over the different
users, shaded areas indicate standard deviations. \andrea{@stefano:
  possiamo riportare la mediana anche per i tempi?}  Our {\sc
  SetMargin} algorithm for $k=2$ is reported in orange, Guo in blue,
Viappiani QI in green and Viappiani EUS in gray. For all algorithms,
one iteration corresponds to a single pairwise query to the user.
Results for $r=3$ and $r=4$ are reported.  For dense weight vector
distributions (first two rows), our approach achieves results which
are indistinguishable from the competitors in a fraction of their
time.  Indeed, all Bayesian approaches cannot scale over $r=4$ (for
$r=5$, Guo did not finish a single user after 24 hours for, Viappiani
QI takes 20 minutes per user), \andrea{20 minuti mi paiono pochi, sei
  sicuro?}  while our algorithm can easily scale to much larger
datasets, as will be shown later on. For sparse weight vector
distributions (last two rows) our approach, in addition to being
substantially faster on each iteration, requires less queries in order
to reach optimal solutions. This is an expected result as the
sparsification norm in our formulation ($\| \vw^{i} \|_1$) is
enforcing sparsity in the weight vectors, while none of the other
approaches are explicitly designed to favour sparse functions.

% Guo e' piantato al primo utente da tre giorni su synthetic 5.
% Paolo procede su S-5 da qualche ora (~8) (1290s su synthetic-5
% uniforme per completamento di un utente).

%\stefano{@Paolo: varianza tempi?}

In order to study the effect of increasing the number of weight
vectors in our formulation, we also ran {\sc SetMargin} varying the
parameter $k$. Figure~\ref{fig:selfcomparison} reports utility loss
results on $r=4$ and $r=5$ datasets for the uniform and normal sparse
distributions\footnote{We picked the most difficult dense distribution
  (uniform) and the simplest sparse distribution (normal) to provide
  the broadest possible picture, as space constraints prevent us from
  including all results. Results for the other distributions are
  however qualitative similar to the ones reported.}, for $k=2$
(orange), $k=3$ (blue) and k=4 (green). The first and third columns
report results in terms of number of iterations. It can be seen that
increasing the number of weight vectors tends to favour earlier
convergence, especially for the more complex dataset ($r=5$). However,
as in each iteration the user is asked to compare $k$ items, different
values of $k$ imply a different cognitive effort for the user. The
second and fourth columns report results in terms of number of
queries, where we count all $k \choose 2$ pairs of queries when
comparing $k$ items. In this case, $k=2$ seems to be the best
option. The cognitive cost for the user will likely lay in between
these two extremes, but formalizing this concept in an efficient query
ordering strategy needs to face the effect of noise.  A modified
sorting algorithm asking only $O(k\log k)$ queries to the user
resulted in a performance worsening, likely because of a cascading
effect of inconsistence feedback. \andrea{@stefano: l'ho rifrasato un
  po', vediamo se ci piace}

% \paragraph{Cognitive effort.} Contrarily to most other preference elicitation
% methods, which generate a single candidate option per iteration, {\sc
% SetMargin} constructs a whole {\em set} of options. We thus can not leverage
% standard query selection strategies \stefano{which ones?}.

% The simplest solution is to query the user about all $k \choose 2$ pairs of
% configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
% number of queries does not grow large in practice.  An appealing alternative
% involves using a modified sorting algorithm, such as merge-sort, to order the
% set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
% approach is very susceptible to noise: a single inconsistent user response may
% affect all successive comparisons, leading to a cascading effect. While more
% robust strategies may be conceived, empirically we found that the simpler
% quadratic scheme tends to work reasonably well.
